The Astropy Project

### 2017-10-23 main.tex
Update on Overleaf.
"IAU SOFA Board" was taken as author name

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
All astronomical research makes use of software in some way and
astronomy as a field has long supported the development of specialized
software tools for different astronomical tasks. Some of those
software packages are (or at least were) supported by large
institutions, e.g.\ IRAF \citep[developed at NOAO,][]{IRAF}, MIDAS
\citep[developed at ESO,][]{MIDAS}, or ds9 \citep[developed at
% * <bsipocz@gmail.com> 2017-09-28T21:05:00.569Z:
%
% > or ds9
% probably starlink is a better example as being a large package of software similarly to IRAF and MIDAS rather than a single tool.
%
% ^.
% hamogu: Feel free ot swap in Starlink, if you think that's more appropriate. I picked ds9 becase (a) it's wider known (e.g. in the X-ray community) and (b) because it is an example of a widely used tool with institutional support that is NOT a large package. TOPCAT would have been another example if we want to expand that list.
  SAO,][]{ds9}. These packages tend to address a wide range of users
and typically provide some level of documentation and user
support is provided. Other packages are developed foremost by individual research
groups and used mostly in work by members of the group. Those
packages can address more specialized needs, but are not necessarily
available to the community and even if they are, documentation can in
some cases be insufficient for new users and the quality of the
software, i.e.\ the robustness and the level of testing that has been
performed to ensure correctness, is not always clear. On the other
hand, there are also many widely used, well documented, high quality
software packages which are developed by individuals or small groups.

@@

\section{Outlook}
\label{sec:outlook}
\inlinecomment{hamogu}{This may be merged into the section above once
  that is fleshed out a little more.}  As shown above the development
of the \astropypkg package is making good progress, helping
astronomers worldwide to perform many daily tasks such as planning
observations, analyzing data or simulation results, and writing
publications. The strong emphasis that the \astropy project puts on
reliability and code correctness helps users to trust the calculations
performed with \astropypkg and to publish reproducable results. More
functionality is added to the core package and more and more
affiliated packages support more specialized needs.

At the same time, \astropy is spreading awareness of best practices in
software development. This is important since most practicing
astronomers received their education without being taught in computer
science and software development although a major fraction of almost
every astronomer's workload today is related to software use and
development. The \astropypkg package leads by example, showing all interested
astronomers how modern tools like git verion control or continuous
integration (code is tested after every change to the source) can
increase the quality and accessability of astronomical software
without overly complicating the development cycle. All submitted code
is reviewed by at least one, but typically more, members of the
astropy community, providing feedback and allowing contributors to
improve their skill is developing software without the need for their
institutions to provide dedicated teaching recources. Thus, the
\astropy project does not only build the tools needed to perform
productive astronomical research in an era of increasingly large and
complex datasets, it also helps to prepare the current and the next
generation of researchers with the knowledge to adequatly use those
tools.

>>>>

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
All astronomical research makes use of software in some way and
astronomy as a field has long supported the development of specialized
software tools for different astronomical tasks. Some of those
software packages are (or at least were) supported by large
institutions, e.g.\ IRAF \citep[developed at NOAO,][]{IRAF}, MIDAS
\citep[developed at ESO,][]{MIDAS}, or ds9 \citep[developed at
% * <bsipocz@gmail.com> 2017-09-28T21:05:00.569Z:
%
% > or ds9
% probably starlink is a better example as being a large package of software similarly to IRAF and MIDAS rather than a single tool.
%
% ^.
% hamogu: Feel free ot swap in Starlink, if you think that's more appropriate. I picked ds9 becase (a) it's wider known (e.g. in the X-ray community) and (b) because it is an example of a widely used tool with institutional support that is NOT a large package. TOPCAT would have been another example if we want to expand that list.
% olebole: TOPCAT is part of Starlink Java. An other examples would be CASA resp. AIPS. I'd rather omit ds9, since it is mainly developed by one person. 
  SAO,][]{ds9}. These packages tend to address a wide range of users
and typically provide some level of documentation and user
support is provided. Other packages are developed foremost by individual research
groups and used mostly in work by members of the group. Those
packages can address more specialized needs, but are not necessarily
available to the community and even if they are, documentation can in
some cases be insufficient for new users and the quality of the
software, i.e.\ the robustness and the level of testing that has been
performed to ensure correctness, is not always clear. On the other
hand, there are also many widely used, well documented, high quality
software packages which are developed by individuals or small groups.

@@

\section{Outlook}
\label{sec:outlook}
\inlinecomment{hamogu}{This may be merged into the section above once
  that is fleshed out a little more.}  As shown above the development
of the \astropypkg package is making good progress, helping
astronomers worldwide to perform many daily tasks such as planning
observations, analyzing data or simulation results, and writing
publications. The strong emphasis that the \astropy project puts on
reliability and code correctness helps users to trust the calculations
performed with \astropypkg and to publish reproducible results. More
functionality is added to the core package and more and more
affiliated packages support more specialized needs.

At the same time, \astropy is spreading awareness of best practices in
software development. This is important since most practicing
astronomers received their education without being taught in computer
science and software development although a major fraction of almost
every astronomer's workload today is related to software use and
development. The \astropypkg package leads by example, showing all interested
astronomers how modern tools like git version control or continuous
integration (code is tested after every change to the source) can
increase the quality and accessibility of astronomical software
without overly complicating the development cycle. All submitted code
is reviewed by at least one, but typically more, members of the
astropy community, providing feedback and allowing contributors to
improve their skill is developing software without the need for their
institutions to provide dedicated teaching resources. Thus, the
\astropy project does not only build the tools needed to perform
productive astronomical research in an era of increasingly large and
complex datasets, it also helps to prepare the current and the next
generation of researchers with the knowledge to adequately use those
tools.
